Principles for Open Data Curation: A Case Study with the New York City 311 Service Request Data

### 2024-06-14 manuscript/manuscript.tex
Added the one JPG file for figure
Deleted older PDF file which wasn't working and added JPG image for computed_fields in NYC Data Portal
Spacing and caption location fixes

The paper is organized as follows:
Section~\ref{sec:history} offers a brief review of the history of the 311 system. In 
Section~\ref{sec:trends} we take a look at long-term trends presented via a 10-year analysis.
Section~\ref{sec:issues} offers a general discussion of data cleansing issues
impacting data quality and curation efficiency. Section~\ref{sec:structural} examines
the technical dimensions of the dataset, the structural issues. In Section~\ref{sec:datatypes} we examine the data fields
for compliance with the stated data type as found in the Data Dictionary. Section~\ref{sec:blanks} looks
at the data fields as regards missing, blank, or N/A entries. Section~\ref{sec:domain} explores
how data fields comply with a domain of legal or acceptable values. Section~\ref{sec:inconsistencies}
deals with the important issues surrounding logical inconsistencies and concerning patterns in the data. Section~\ref{sec:precision} 
explores the age-old issue of precision versus accuracy. Section~\ref{sec:duplicates}
identifies duplicate and redundant data fields. Section~\ref{sec:improvements} provides 
actionable suggestions for mitigating or resolving identified issues. Following this, Section~\ref{sec:protocol} outlines a series of general 
principles for the release of open data, drawing from our findings. The paper concludes with a discussion 
in Section~\ref{sec:discussion}, encapsulating the key insights and implications of our research.

@@

\begin{table}[tbp]
    \centering
    \caption{Noise-related complaints\_type(s) by count with Agency}
    \normalsize
    \begin{tabular}{@{}lS[table-format=7.0,round-mode=places,round-precision=0]S[table-format=2.2,round-mode=places,round-precision=2]l@{}} % 'l' for left-aligned, 'S' for siunitx number alignment
        \toprule
        \textbf{complaint\_type} & \textbf{Count} & \textbf{Percentage} & \textbf{Agency} \\ 
        \midrule
        NOISE - RESIDENTIAL        & 675502 & 10.56 & NYPD  \\ 
        NOISE - STREET/SIDEWALK    & 300507 &  4.70 & NYPD  \\ 
        NOISE - COMMERCIAL         & 128892 &  2.02 & NYPD  \\ 
        NOISE - VEHICLE            & 115956 &  1.81 & NYPD  \\ 
        NOISE                      & 100413 &  1.57 & DEP   \\ 
        NOISE - HELICOPTER         &  85345 &  1.33 & EDC   \\ 
        NOISE - PARK               &  16620 &  0.26 & NYPD  \\ 
        NOISE - HOUSE OF WORSHIP   &   2757 &  0.04 & NYPD  \\ 
        \bottomrule
    \end{tabular}
    \label{tab:noisecomplaints}
\end{table}

@@

\textbf{Issue: Fields not in the Data Dictionary} The 311 SR \href{https://data.cityofnewyork.us/api/views/erm2-nwe9/files/b372b884-f86a-453b-ba16-1fe06ce9d212?download=true&filename=311_ServiceRequest_2010-Present_DataDictionary_Updated_2023.xlsx}{Data Dictionary}
 identifies 41 data columns (fields) along with related information for each column. 
 However, when you download the data or explore it via the online portal, you immediately note that there are 47 columns (fields).
 he additional six fields are not addressed in the Data Dictionary, but do and show up on the Column Manager widget
 as ``@computed\_region\_xxxx\_xxxx''. To some extent one can infer what the these @computed\_fields are, but it is not possible to know
for certain. These six fields are:

@@

\begin{figure}[tbp]
  \centering
 	\includegraphics[width = \textwidth] {computed_fields.pdf}	
	\caption{Screenshot of Data Portal with ``computed'' Columns}  
	\label{fig:computed-columns}
\end{figure}

@@

\begin{itemize}
	\item created\_date, closed\_date, due\_date, and resolution\_action\_updated\_date: All are all in proper date format.
	\item zip\_codes and incident\_zip: All are 5 numeric digits except for 2 non-numeric entries for incident\_zip (``na'' and ``N/A'').
	\item x\_coordinate\_state\_plane \&  y\_coordinate\_state\_plane (a NY State geo-location system): All are all numeric.
	\item latitude \& longitude: Both are numeric.
	\item community\_districts, borough\_boundaries, city\_council\_district, police\_precinct, and police\_precincts: All are numeric.
	\item Free-form text fields that contain a ``,'' (e.g. incident\_address \& resolution\_description) are properly enclosed in quotation marks.
\end{itemize}

All data fields were subjected to a test of ``correctness'' according to their ``type'' as specified
in the Data Dictionary. Of the nearly 300 million data fields to check (6.5 million rows, 47 fields per row), there were only two values that failed to test the
test for appropriate data type; those being two text entries in the incident\_zip field (``na'', and ``N/A''). All date fields were dates. All numbers were numeric.
All text fields were character. Data typing appears to not be an issue with this dataset.

@@

When counting the various fields for blank or N/A values, the fields appear to divide into three groups: \textbf{Mostly Empty}, \textbf{Partially Empty}, 
and \textbf{Few/None Empty}. he Mostly Empty category ranges from 93-99.9\% blank. It includes such fields as taxi\_company due\_date, pickup\_location, and landmark. The Partially 
Empty includes such fields as location\_type, borough, and cross\_street. And the Few/None Empty includes created\_date, complaint\_type, agency, and status.
In some cases it may make sense to inquiry as to why some fields are frequently blank. Is the data difficult to capture? Does it pertain to only a small set
of complaint\_type(s) or Agencies? And if the data is truly that sparse, does it even make sense to collect it? Is it legacy data? For example, one would expect that every complaint\_type
that has a geographical component, such as an address, would also have a corresponding police\_precinct, community\_board, and community\_district since these
are all-inclusive throughout the five Boroughs of NYC. Addtionally, that it would have a lat/long. And generally that is the case, but these fields do have small number, but nonetheless significant, 
of empty entries.

@@

\begin{table}[tbp]
\centering
\caption{Blank and N/A entries by Field}
\resizebox{0.8\linewidth}{!}{%
\normalsize
\begin{tabular}{lrrrl}
	\toprule
	\textbf{Field} & \textbf{Total Empty} & \textbf{Percent Empty} & \textbf{N/A Count} \\
	\midrule
		taxi\_company\_borough & 6391341 & 99.94 & 0 \\
		road\_ramp & 6379080 & 99.75 & 0 \\
		vehicle\_type & 6376266 & 99.71 & 0 \\
		due\_date & 6370497 & 99.62 & 6370497 \\
		bridge\_highway\_direction & 6367626 & 99.57 & 0 \\
		bridge\_highway\_name & 6340197 & 99.14 & 0 \\
		bridge\_highway\_segment & 6340190 & 99.14 & 0 \\
		taxi\_pick\_up\_location & 6329954 & 98.98 & 0 \\
		facility\_type & 5966588 & 93.30 & 0 \\
		landmark & 2635736 & 41.22 & 0 \\
		intersection\_street\_1 & 2143847 & 33.52 & 0 \\
		intersection\_street\_2 & 2140707 & 33.48 & 0 \\
		cross\_street\_1 & 1846594 & 28.88 & 0 \\
		cross\_street\_2 & 1846008 & 28.87 & 0 \\
		location\_type & 798775 & 12.49 & 0 \\
		bbl & 768109 & 12.01 & 0 \\
		city & 332086 & 5.19 & 0 \\
		street\_name & 280463 & 4.39 & 0 \\
		incident\_address & 280268 & 4.38 & 0 \\
		closed\_date & 248839 & 3.89 & 248839 \\
		resolution\_description & 132333 & 2.07 & 0 \\
		zip\_codes & 123167 & 1.93 & 0 \\
		borough\_boundaries & 101453 & 1.59 & 0 \\
		community\_districts & 101445 & 1.59 & 0 \\
		city\_council\_districts & 101445 & 1.59 & 0 \\
		police\_precincts & 101445 & 1.59 & 0 \\
		police\_precinct & 101416 & 1.59 & 0 \\
		latitude & 99778 & 1.56 & 0 \\
		longitude & 99778 & 1.56 & 0 \\
		location & 99778 & 1.56 & 0 \\
		x\_coordinate\_state\_plane & 99668 & 1.56 & 0 \\
		y\_coordinate\_state\_plane & 98822 & 1.55 & 0 \\
		resolution\_action\_updated\_date & 91399 & 1.43 & 91399 \\
		incident\_zip & 83220 & 1.30 & 2 \\
		descriptor & 74645 & 1.17 & 0 \\
		address\_type & 38655 & 0.60 & 0 \\
		unique\_key & 0 & 0.00 & 0 \\
		agency & 0 & 0.00 & 0 \\
		agency\_name & 0 & 0.00 & 0 \\
		complaint\_type & 0 & 0.00 & 0 \\
		status & 0 & 0.00 & 0 \\
		community\_board & 0 & 0.00 & 0 \\
		borough & 0 & 0.00 & 0 \\
		open\_data\_channel\_type & 0 & 0.00 & 0 \\
		park\_facility\_name & 0 & 0.00 & 0 \\
		park\_borough & 0 & 0.00 & 0 \\
		created\_date & 0 & 0.00 & 0 \\
	\bottomrule
\end{tabular}
}
\end{table}

@@

\begin{figure}[tbp]
  \centering
  	\caption{Number and Percentage of Empty/Blank Entries}
	\includegraphics[width=\textwidth]{BlankFields.pdf}
	\label{fig:blank_fields}
\end{figure}

@@

	Furthermore, the field incident\_zip while having only .07\% invalid entries, still is a problem with 4163 such errors.
	As the zip code is a primary measure of many NYC city services, 
	and with a freely available database for which to validate entries, it seems an oversight 
	that any such errors could creep into the 311 system. Some invalid entries can  
	clearly be identified as incorrect just by observation, e.g. 10000, 12345, 11111, etc.

@@

	The  errors in the incident\_zip field are more troubling, even though they are, percentage-wise, small. Here is a graphic 
	illustrating how these errors occur by Agency. Here we can see that the majority of these errors lie in the Dept of Transportation,
	NYPD, Taxi \& Limo Commission (TLC), and the Economic Development Council (EDC) among others. This representation
	does not follow the full SR distribution indicating that these errors are likely generated by an incorrect
	process or application residing at those particular Agencies; it is not a systemic problem.

	\begin{figure}[tbp]
		\centering
		\caption{Invalid incident\_zip(s) by Agency}
		\includegraphics[width=\textwidth]{invalid_incident_zip.pdf} 
		\label{fig:invalid_incident_zip}
	\end{figure}

		\subsubsection{Case Study: Noise Complaints by Zip Code}{label{sec:case-study-zip-codes}
		\textbf{Scenario:} The NYC Office of Nightlife (ONL) wants to know ``What are the top 10 zip codes for Noise Complaints (all 8 types) over the last two years?''
		The goal is to assess the impact of the recent NYC effort looking to promote a safe and vibrant nightlife scene in NYC while seeking to ease
		strained relations between bar and club owners.

@@

\begin{table}[tbp]
    \centering
    \normalsize
    \caption{Comparison of Top Ten Zip Codes Lists}
    \begin{tabular}{@{}lS[table-format=6.0,round-mode=places,round-precision=0]c@{\hskip 0.5cm}@{}lS[table-format=6.0,round-mode=places,round-precision=0]c@{}}
        \toprule
        \multicolumn{3}{c}{\textbf{zip\_codes}} & \multicolumn{3}{c}{\textbf{incident\_zip}} \\
        \cmidrule(r){1-3} \cmidrule(l){4-6}
        \textbf{Zip Code} & \textbf{Count} & \textbf{Valid?} & \textbf{Zip Code} & \textbf{Count} & \textbf{Valid?} \\
        \midrule
        11275 & 104556 & FALSE & 10466 & 104562 & TRUE \\
        12420 & 27503 & TRUE & 10023 & 27972 & TRUE \\
        12428 & 26564 & TRUE & 10031 & 25548 & TRUE \\
        10935 & 25508 & FALSE & 10457 & 25066 & TRUE \\
        10934 & 23448 & FALSE & 10453 & 24752 & TRUE \\
        10931 & 22381 & TRUE & 10456 & 24751 & TRUE \\
        10930 & 22121 & TRUE & 10452 & 22527 & TRUE \\
        17613 & 21963 & FALSE & 10025 & 21705 & TRUE \\
        10936 & 21707 & FALSE & 10458 & 21689 & TRUE \\
        11606 & 21435 & FALSE & 10032 & 20622 & TRUE \\
        \bottomrule
    \end{tabular}
    \label{tab:zipcodes}
\end{table}

@@

	\subsection{Issues with Police Precincts}
	\label{sec:police-precincts}
	A curious case also exists when examining the two nearly identical fields - police\_precincts and police\_precinct. Both of those fields are among
	the ``computed'' fields in the dataset. 
	Using \href{https://www.nyc.gov/site/nypd/bureaus/patrol/precincts-landing.page}{NYPD Precinct listings} it's possible to determine the valid
	police precincts; they are generally numeric or have numeric representations in the dataset.

@@

	For the police\_precincts field, there are 2,171,864 invalid entries (35\%)., a consequential number of errors. Similarly for the police\_precinct field,
	 there are 2,171,778 invalid entries ( also 35\%), however they are not the same counts for the two similar fields.

@@

	\begin{figure}[tbp]
	  \centering
	  \caption{Invalid NYPD precincts by Agency}
	  \includegraphics[width=\textwidth]{invalid_police_precincts.pdf}	  
	  \label{fig:invalid_police_precincts_zip}
	\end{figure}

	\subsection{Issues with Community Boards}
	Community Boards are an important aspect of NYC government. They are the most local, grassroots form of City government, 
	and a vital connection between communities and elected officials and City agencies. They are organized by the five Boroughs of NYC and represented
	in this dataset as ``\#\#-Borough'', e.g. ``10 Bronx'' and report to the various Borough Presidents in NYC. As such, they play an important role in the quality of life for all New Yorkers.
	There are 59 community boards throughout the City: 12 Community Districts in the Bronx, 18 in Brooklyn, 12 in Manhattan, 14 in Queens and 3 in Staten Island.
	The Community Boards are used frequently as a way to measure distribution of City services throughout the five boroughs, so correctness is important. 
	NYC Borough Presidents in particular are focused on the services offered at the Community Board level within their respective Borough.

@@

		\begin{table}[tbp]
		\centering
		\caption{Top Ten Invalid 'community\_board' Values}
		\normalsize
		\begin{tabular}{rlr}
		\toprule
		\textbf{Rank} & \textbf{Invalid CB} & \textbf{Count} \\
			\midrule
				1 & 64 MANHATTAN & 7144 \\
				2 & 83 QUEENS & 5133 \\
				3 & 55 BROOKLYN & 3327 \\
				4 & 81 QUEENS & 3180 \\
				5 & 80 QUEENS & 2407 \\
				6 & 26 BRONX & 2364 \\
				7 & 28 BRONX & 1128 \\
				8 & 82 QUEENS & 1054 \\
				9 & 95 STATEN ISLAND & 515 \\
				10 & 27 BRONX & 514 \\
			\bottomrule
		\end{tabular}
		\end{table}

@@

\subsection{created\_date and closed\_date(s) -- Negative Duration}
Let's look first at the created\_date and closed\_date fields. You might expect that these two fields are automatically populated by the 
SR application software, either the software used at the 311 organization or that used at the Agencies that are integrated with the 311 system;
and typically that is indeed the case. In such a case, the act of creating an SR would automatically trigger an entry in created\_date based on a system-wide clock. Thus
the created\_date is associated with setting an SR status to ``new'' or ``open''. And similarly when an SR's status is changed to
``closed'' then the closed\_date would be populated.

Unfortunately, our analysis found that that assumption is not always the case. When you analyze the various date fields, several
anomalies become quickly apparent:

@@

\begin{table}[tbp]
    \centering
    \caption{Largest and Smallest errors (days)}
    \normalsize
    \begin{tabular}{>{\ttfamily}l l l r l}
        \toprule
        \multicolumn{5}{c}{\textbf{Largest errors (days) excluding extreme negative values}} \\
        \midrule
        \textbf{created\_date} & \textbf{closed\_date} & \textbf{duration} & \textbf{agency} \\
	        \midrule
	        2023-01-27 14:40:00 & 2022-01-14 14:40:00 & -378.0 & DOT \\
	        2023-01-18 10:06:00 & 2022-01-12 10:06:00 & -371.0 & DOT \\
	        2023-01-27 14:36:00 & 2022-01-22 14:35:00 & -370.0 & DOT \\
	        2023-01-11 11:10:00 & 2022-01-09 11:10:00 & -367.0 & DOT \\
	        2023-12-18 03:13:00 & 2023-01-16 13:10:00 & -335.6 & DOT \\
	        \midrule
	        \multicolumn{5}{c}{\textbf{Smallest errors (days)}} \\
	        \midrule
	        \textbf{created\_date} & \textbf{closed\_date} & \textbf{duration} & \textbf{agency} \\
	        \midrule
	        2023-06-28 07:07:31 & 2023-06-28 07:07:00 & -0.00036 & DOT \\
	        2023-06-29 09:10:20 & 2023-06-29 09:10:00 & -0.00023 & DOT \\
	        2023-01-12 06:50:13 & 2023-01-12 06:50:00 & -0.00015 & DOT \\
	        2023-06-26 08:09:07 & 2023-06-26 08:09:00 & -0.00008 & DOT \\
	        2023-01-12 06:51:01 & 2023-01-12 06:51:00 & -0.00001 & DOT \\
	        \bottomrule
    \end{tabular}
    \label{tab:combined_errors}
\end{table}

@@

According, these  SR rows are removed from the box \& whiskers plot. Note that these extremely large negative duration SRs all
are from the Department of Homeland Services (DHS) Sample:

>>>>

The paper is organized as follows:
Section~\ref{sec:history} offers a brief review of the history of the 311 system. In 
Section~\ref{sec:trends} we take a look at long-term trends presented via a 10-year analysis.
Section~\ref{sec:issues} offers a general discussion of data cleansing issues
impacting data quality and curation efficiency. Section~\ref{sec:structural} examines
the technical dimensions of the dataset, the structural issues. 
In Section~\ref{sec:datatypes} we examine the data fields for compliance 
with the stated data type as found in the Data Dictionary. Section~\ref{sec:blanks} looks
at the data fields as regards missing, blank, or N/A entries. Section~\ref{sec:domain} explores
how data fields comply with a domain of legal or acceptable values. Section~\ref{sec:inconsistencies}
deals with the important issues surrounding logical inconsistencies 
and concerning patterns in the data. Section~\ref{sec:precision} 
explores the age-old issue of precision versus accuracy. Section~\ref{sec:duplicates}
identifies duplicate and redundant data fields. Section~\ref{sec:improvements} provides 
actionable suggestions for mitigating or resolving identified issues. Following this, Section~\ref{sec:protocol} outlines a series of general 
principles for the release of open data, drawing from our findings. 
The paper concludes with a discussion and recommendations
in Section~\ref{sec:discussion}, encapsulating the key insights and implications of our research.

@@

\begin{table}[tbp]
    \centering
    \normalsize
    \caption{Noise-related complaints\_type(s) by count with Agency}
    \begin{tabular}{@{}lS[table-format=7.0,round-mode=places,
    round-precision=0]S[table-format=2.2,round-mode=places,
    round-precision=2]l@{}} % 'l' for left-aligned, 'S' for siunitx number alignment
        \toprule
        \textbf{complaint\_type} & \textbf{Count} & \textbf{Percentage} & \textbf{Agency} \\ 
        \midrule
        NOISE - RESIDENTIAL        & 675502 & 10.56 & NYPD  \\ 
        NOISE - STREET/SIDEWALK    & 300507 &  4.70 & NYPD  \\ 
        NOISE - COMMERCIAL         & 128892 &  2.02 & NYPD  \\ 
        NOISE - VEHICLE            & 115956 &  1.81 & NYPD  \\ 
        NOISE                      & 100413 &  1.57 & DEP   \\ 
        NOISE - HELICOPTER         &  85345 &  1.33 & EDC   \\ 
        NOISE - PARK               &  16620 &  0.26 & NYPD  \\ 
        NOISE - HOUSE OF WORSHIP   &   2757 &  0.04 & NYPD  \\ 
        \bottomrule
    \end{tabular}
    \label{tab:noisecomplaints}
\end{table}

@@

\textbf{Issue: Fields not in the Data Dictionary} The 311 SR 
\href{https://data.cityofnewyork.us/api/views/erm2-nwe9/files/b372b884-f86a-453b-ba16-1fe06ce9d212?download=true&filename=311_ServiceRequest_2010-Present_DataDictionary_Updated_2023.xlsx}{Data Dictionary}
 identifies 41 data columns (fields) along with related information 
 for each column. However, when you download the data or 
 explore it via the online portal, you immediately note that there 
 are 47 columns (fields). The additional six fields are not addressed in the 
 Data Dictionary, but do and show up on the Column Manager widget
 as ``@computed\_region\_xxxx\_xxxx''. To some extent one can infer 
 what the these @computed\_fields are, but it is not possible to know 
 for certain. These six fields are:

@@

\begin{figure}[tbp]
  \centering
 	\includegraphics[width = .65\textwidth] {computed_fields_jpg.jpg}	
	\caption{Screenshot of Data Portal with ``@computed'' Columns}  
	\label{fig:computed-columns}
\end{figure}

@@

\begin{itemize}
	\item created\_date, closed\_date, due\_date, and resolution\_action\_updated\_date: 
	All are all in proper date format.
	\item zip\_codes and incident\_zip: All are 5 numeric digits except 
	for two non-numeric entries for incident\_zip (``na'' and ``N/A'').
	\item x\_coordinate\_state\_plane \&  y\_coordinate\_state\_plane 
	(a NY State geo-location system): All are all numeric.
	\item latitude \& longitude: Both are numeric.
	\item community\_districts, borough\_boundaries, 
	city\_council\_district, police\_precinct, and police\_precincts: All are numeric.
	\item Free-form text fields that contain a ``,'' (e.g. incident\_address 
	\& resolution\_description) are properly enclosed in quotation marks.
\end{itemize}

All data fields were subjected to a test of ``correctness'' 
according to their ``type'' as specified
in the Data Dictionary. Of the nearly 300 million data fields to check 
(6.5 million rows, 47 fields per row), there were only two values that failed 
the test for appropriate data type; those being two text entries in the 
incident\_zip field (``na'', and ``N/A''). All date fields were dates. 
All numbers were numeric. All text fields were character. 
Data typing appears to not be an issue with this dataset.

@@

When counting the various fields for blank or N/A values, the fields appear 
to divide into three groups: \textbf{Mostly Empty}, \textbf{Partially Empty}, 
and \textbf{Few/None Empty}. he Mostly Empty category ranges from 93-99.9\% blank. 
It includes such fields as taxi\_company due\_date, pickup\_location, and 
landmark. The Partially Empty includes such fields as location\_type, borough, 
and cross\_street. And the Few/None Empty includes created\_date, 
complaint\_type, agency, and status. In some cases it may make 
sense to inquiry as to why some fields are frequently blank. Is the data 
difficult to capture? Does it pertain to only a small set of 
complaint\_type(s) or Agencies? And if the data is truly that sparse, 
does it make sense to collect it? Is it legacy data? For example, 
one would expect that every complaint\_type that has a geographical 
component, such as an address, would also have a corresponding 
police\_precinct, community\_board, and community\_district since these
are all-inclusive throughout the five Boroughs of NYC. Addtionally, 
that it would have a lat/long. And generally that is the case, but these fields 
do have small number, but nonetheless significant, of empty entries.

@@

\begin{table}[tbp]
\centering
\normalsize
\caption{Blank and N/A entries by Field}
\resizebox{0.8\linewidth}{!}{%
\begin{tabular}{lrrrl}
	\toprule
	\textbf{Field} & \textbf{Total Empty} & \textbf{Percent Empty} 
	& \textbf{N/A Count} \\
	\midrule
		taxi\_company\_borough & 6391341 & 99.94 & 0 \\
		road\_ramp & 6379080 & 99.75 & 0 \\
		vehicle\_type & 6376266 & 99.71 & 0 \\
		due\_date & 6370497 & 99.62 & 6370497 \\
		bridge\_highway\_direction & 6367626 & 99.57 & 0 \\
		bridge\_highway\_name & 6340197 & 99.14 & 0 \\
		bridge\_highway\_segment & 6340190 & 99.14 & 0 \\
		taxi\_pick\_up\_location & 6329954 & 98.98 & 0 \\
		facility\_type & 5966588 & 93.30 & 0 \\
		landmark & 2635736 & 41.22 & 0 \\
		intersection\_street\_1 & 2143847 & 33.52 & 0 \\
		intersection\_street\_2 & 2140707 & 33.48 & 0 \\
		cross\_street\_1 & 1846594 & 28.88 & 0 \\
		cross\_street\_2 & 1846008 & 28.87 & 0 \\
		location\_type & 798775 & 12.49 & 0 \\
		bbl & 768109 & 12.01 & 0 \\
		city & 332086 & 5.19 & 0 \\
		street\_name & 280463 & 4.39 & 0 \\
		incident\_address & 280268 & 4.38 & 0 \\
		closed\_date & 248839 & 3.89 & 248839 \\
		resolution\_description & 132333 & 2.07 & 0 \\
		zip\_codes & 123167 & 1.93 & 0 \\
		borough\_boundaries & 101453 & 1.59 & 0 \\
		community\_districts & 101445 & 1.59 & 0 \\
		city\_council\_districts & 101445 & 1.59 & 0 \\
		police\_precincts & 101445 & 1.59 & 0 \\
		police\_precinct & 101416 & 1.59 & 0 \\
		latitude & 99778 & 1.56 & 0 \\
		longitude & 99778 & 1.56 & 0 \\
		location & 99778 & 1.56 & 0 \\
		x\_coordinate\_state\_plane & 99668 & 1.56 & 0 \\
		y\_coordinate\_state\_plane & 98822 & 1.55 & 0 \\
		resolution\_action\_updated\_date & 91399 & 1.43 & 91399 \\
		incident\_zip & 83220 & 1.30 & 2 \\
		descriptor & 74645 & 1.17 & 0 \\
		address\_type & 38655 & 0.60 & 0 \\
		unique\_key & 0 & 0.00 & 0 \\
		agency & 0 & 0.00 & 0 \\
		agency\_name & 0 & 0.00 & 0 \\
		complaint\_type & 0 & 0.00 & 0 \\
		status & 0 & 0.00 & 0 \\
		community\_board & 0 & 0.00 & 0 \\
		borough & 0 & 0.00 & 0 \\
		open\_data\_channel\_type & 0 & 0.00 & 0 \\
		park\_facility\_name & 0 & 0.00 & 0 \\
		park\_borough & 0 & 0.00 & 0 \\
		created\_date & 0 & 0.00 & 0 \\
	\bottomrule
\end{tabular}
}
\end{table}

@@

\begin{figure}[tbp]
	\centering
  	\includegraphics[width=\textwidth]{BlankFields.pdf}
	\caption{Number and Percentage of Empty/Blank Entries}
	\label{fig:blank_fields}
\end{figure}

@@

	Furthermore, the field incident\_zip while having only .07\% invalid entries, 
	still is a problem with 4163 such errors. As the zip code is a primary 
	measure of many NYC city services, and with a freely available 
	database for which to validate entries, it seems an oversight 
	that any such errors could creep into the 311 system. 
	Some invalid entries can clearly be identified as incorrect by 
	observation, e.g. 10000, 12345, 11111, etc.

@@

	The  errors in the incident\_zip field are more troubling, even though 
	they are, percentage-wise, small. Here is a graphic illustrating how 
	these errors occur by Agency. Here we can see that the majority of 
	the errors lie in the Dept of Transportation, NYPD, Taxi \& Limo 
	Commission (TLC), and the Economic Development Council (EDC) 
	among others. This representation does not follow the full SR 
	distribution indicating that these errors are likely generated by 
	an incorrect process or application residing at those particular 
	Agencies; it is not a systemic problem.

	\begin{figure}[tbp]
		\centering
		\includegraphics[width=\textwidth]{invalid_incident_zip.pdf}
		\caption{Invalid incident\_zip(s) by Agency} 
		\label{fig:invalid_incident_zip}
	\end{figure}

		\subsubsection{Case Study: Noise Complaints by Zip Code}
		{label{sec:case-study-zip-codes}
		\textbf{Scenario:} The NYC Office of Nightlife (ONL) wants to 
		know ``What are the top 10 zip codes for Noise Complaints (all 8 types) 
		over the last two years?'' The goal is to assess the impact of the 
		recent NYC effort looking to promote a safe and vibrant 
		nightlife scene in NYC while seeking to ease strained 
		relations between bar and club owners.

@@

\begin{table}[tbp]
    \centering
    \normalsize
    \caption{Comparison of Top Ten Zip Codes Lists}
    \begin{tabular}{@{}lS[table-format=6.0,round-mode=places,
    round-precision=0]c@{\hskip 0.5cm}@{}lS[table-format=6.0,
    round-mode=places,round-precision=0]c@{}}
        \toprule
        \multicolumn{3}{c}{\textbf{zip\_codes}} & \multicolumn{3}{c}{\textbf{incident\_zip}} \\
        \cmidrule(r){1-3} \cmidrule(l){4-6}
        \textbf{Zip Code} & \textbf{Count} & \textbf{Valid?} 
        & \textbf{Zip Code} & \textbf{Count} & \textbf{Valid?} \\
        \midrule
        11275 & 104556 & FALSE & 10466 & 104562 & TRUE \\
        12420 & 27503 & TRUE & 10023 & 27972 & TRUE \\
        12428 & 26564 & TRUE & 10031 & 25548 & TRUE \\
        10935 & 25508 & FALSE & 10457 & 25066 & TRUE \\
        10934 & 23448 & FALSE & 10453 & 24752 & TRUE \\
        10931 & 22381 & TRUE & 10456 & 24751 & TRUE \\
        10930 & 22121 & TRUE & 10452 & 22527 & TRUE \\
        17613 & 21963 & FALSE & 10025 & 21705 & TRUE \\
        10936 & 21707 & FALSE & 10458 & 21689 & TRUE \\
        11606 & 21435 & FALSE & 10032 & 20622 & TRUE \\
        \bottomrule
    \end{tabular}
    \label{tab:zipcodes}
\end{table}

@@

	\subsection{Issues with Police Precincts}
	\label{sec:police-precincts}
	A curious case also exists when examining the two nearly identical 
	fields - police\_precincts and police\_precinct. Both of those fields 
	are among the ``computed'' fields in the dataset. Using 
	\href{https://www.nyc.gov/site/nypd/bureaus/patrol/precincts-landing.page}
	{NYPD Precinct listings} it's possible to determine the valid 
	police precincts; they are generally numeric or have numeric 
	representations in the dataset.

@@

	For the police\_precincts field, there are 2,171,864 invalid entries 
	(35\%)., a consequential number of errors. Similarly for the 
	police\_precinct field, there are 2,171,778 invalid entries 
	(also 35\%), however they are not the same counts for the two similar fields.

@@

	\begin{figure}[tbp]
	  \centering
	  \includegraphics[width=\textwidth]{invalid_police_precincts.pdf}
	  \caption{Invalid NYPD precincts by Agency}	  
	  \label{fig:invalid_police_precincts_zip}
	\end{figure}

	\subsection{Issues with Community Boards}
	Community Boards are an important aspect of NYC government. They are the 
	most local, grassroots form of City government, and a vital connection 
	between communities and elected officials and City agencies. 
	They are organized by the five Boroughs of NYC and represented	
	in this dataset as ``\#\#-Borough'', e.g. ``10 Bronx'' and report to the 
	various Borough Presidents in NYC. As such, they play an important 
	role in the quality of life for all New Yorkers. There are 59 community 
	boards throughout the City: 12 Community Districts in Bronx, 18 in Brooklyn, 
	12 in Manhattan, 14 in Queens and 3 in Staten Island. The Community 
	Boards are used frequently as a way to measure distribution of 
	City services throughout the five boroughs, so correctness is important. 
	NYC Borough Presidents in particular are focused on the services 
	offered at the Community Board level within their respective Borough.

@@

		\begin{table}[tbp]
		\centering
		\normalsize
		\caption{Top Ten Invalid 'community\_board' Values}
		\begin{tabular}{rlr}
		\toprule
		\textbf{Rank} & \textbf{Invalid CB} & \textbf{Count} \\
			\midrule
				1 & 64 MANHATTAN & 7144 \\
				2 & 83 QUEENS & 5133 \\
				3 & 55 BROOKLYN & 3327 \\
				4 & 81 QUEENS & 3180 \\
				5 & 80 QUEENS & 2407 \\
				6 & 26 BRONX & 2364 \\
				7 & 28 BRONX & 1128 \\
				8 & 82 QUEENS & 1054 \\
				9 & 95 STATEN ISLAND & 515 \\
				10 & 27 BRONX & 514 \\
			\bottomrule
		\end{tabular}
		\end{table}

@@

\subsection{created\_date and closed\_date(s) -- Negative Duration}
Let's look first at the created\_date and closed\_date fields. You might 
expect that these two fields are automatically populated by the 
SR application software, either the software used at the 311 organization 
or that used at the Agencies that are integrated with the 311 system; 
and typically that is indeed the case. In such a case, the act of creating an 
SR would automatically trigger an entry in created\_date based on a 
system reference clock. Thus the created\_date is associated with 
setting an SR status to ``new'' or ``open''. And similarly when an SR's 
status is changed to ``closed'' then the closed\_date would be populated.

Unfortunately, our analysis found that that assumption is not always 
the case. When you analyze the various date fields, several anomalies 
quickly become apparent:

@@

\begin{table}[tbp]
    \centering
    \normalsize
    \caption{Largest and Smallest errors (days)}
    \begin{tabular}{>{\ttfamily}l l l r l}
        \toprule
        \multicolumn{5}{c}{\textbf{Largest errors (days) excluding 
        extreme negative values}} \\
        \midrule
        \textbf{created\_date} & \textbf{closed\_date} & \textbf{duration} 
        & \textbf{agency} \\
	        \midrule
	        2023-01-27 14:40:00 & 2022-01-14 14:40:00 & -378.0 & DOT \\
	        2023-01-18 10:06:00 & 2022-01-12 10:06:00 & -371.0 & DOT \\
	        2023-01-27 14:36:00 & 2022-01-22 14:35:00 & -370.0 & DOT \\
	        2023-01-11 11:10:00 & 2022-01-09 11:10:00 & -367.0 & DOT \\
	        2023-12-18 03:13:00 & 2023-01-16 13:10:00 & -335.6 & DOT \\
	        \midrule
	        \multicolumn{5}{c}{\textbf{Smallest errors (days)}} \\
	        \midrule
	        \textbf{created\_date} & \textbf{closed\_date} & \textbf{duration} 
	        & \textbf{agency} \\
	        \midrule
	        2023-06-28 07:07:31 & 2023-06-28 07:07:00 & -0.00036 & DOT \\
	        2023-06-29 09:10:20 & 2023-06-29 09:10:00 & -0.00023 & DOT \\
	        2023-01-12 06:50:13 & 2023-01-12 06:50:00 & -0.00015 & DOT \\
	        2023-06-26 08:09:07 & 2023-06-26 08:09:00 & -0.00008 & DOT \\
	        2023-01-12 06:51:01 & 2023-01-12 06:51:00 & -0.00001 & DOT \\
	        \bottomrule
    \end{tabular}
    \label{tab:combined_errors}
\end{table}

@@

According, these  SR rows are removed from the box \& whiskers plot. 
Note that these extremely large negative duration SRs are all from the 
Department of Homeland Services (DHS) Sample:

\begin{table}[tbp]
    \centering
     \normalsize
    \caption{Sample of SRs with extremely large negative durations}
    \begin{tabular}{l l r l}
        \toprule
        \textbf{created\_date} & \textbf{closed\_date} & \textbf{duration} 
        & \textbf{agency} \\
	        \midrule
	        2022-02-11 13:43:05 & 1900-01-01 & -44602 & DHS \\
	        2022-02-11 11:03:38 & 1900-01-01 & -44602 & DHS \\
	        2022-02-11 08:29:49 & 1900-01-01 & -44601 & DHS \\
	        2022-02-11 14:08:08 & 1900-01-01 & -44602 & DHS \\
	        2022-02-11 14:05:16 & 1900-01-01 & -44602 & DHS \\
	        \bottomrule
    \end{tabular}
    \label{tab:extreme_negative_durations}
\end{table}
